### 2013-07-24 3/sar.tex
updated for Ian changes

\chapter{Consistent Measures of Distance for Univariate SAR Data} %chapter 3
\label{chap:sar}

%\section{SAR speckle statistic model}
%\label{sec:heteroskdastic_model}

Useful and robust statistical estimators require un-assuming models. 
In this chapter a homoskedastic model for the usually heteroskedastic SAR is proposed. %speckle filtering.
In the first section the heteroskedastic nature of original amplitude and intensity SAR data will be explained. 
Relying solely on statistical principles,
  log-transformed SAR data is shown to be homoskedastic.
Together with the consistent sense of variance, a few consistent measures of distance are proposed for the traditional SAR data.

@@

SAR speckle phenomena are explained as the interference of many coherent but dephased back-scaterring components;
  each reflecting from different and distributed elementary scaterers \citep{Oliver_ProcIEEE_1963, Leith_ProcIEEE_1971}. 
The random nature of the SAR data arises
  due to the unknown and varying location, height and thus distance of each elementary scaterer;
  which, in turn, generates random phases in their back-scattering responses.
Assuming the number of these elementary back scatterers is sufficiently large,
  then the Central Limit Theorem is applicable \citep{Goodman_Springer_1975}.
Consequently the interference of these randomly phased responses can be considered as
  a random walk on the 2D complex plane \citep{Goodman_JOptSocAm_76}.  
Specifically the real part $A_r$ as well as the imaginary part $A_i$ of the observed SAR signal $A$ can be considered as
  random variables from uncorrelated Gaussian distributed stochastic processes with zero means and indentical variances $\sigma^2/2$  \citep{Lee_CRCPress_2009}. 
Their probability density functions (PDF) are given as:

\begin{equation}
\label{eqn:SAR Real and Imaginary Components PDF}
\caption{eqn:SAR Real and Imaginary Components PDF}
pdf(A_x)=\frac{1}{\sqrt{\pi} \sigma} e^{\left( \frac{A_x^2}{\sigma^2} \right) }
\end{equation}

It can then be proved that the measurable amplitude $A=\sqrt{A_r^2+A_i^2}$ is a random variable of Rayleigh distribution
  and subsequently the SAR intensity $I=A^2=(A_r^2+A_i^2)$ is a random variable of a negative exponentially distributed random process.
That is:

\begin{eqnarray}
pdf(A) &=& \frac{2A}{\sigma^2}e^{ \left( -\frac{A^2}{\sigma^2} \right) }\\
pdf(I) &=& \frac{1}{\sigma^2}e^{\left( -\frac{I}{\sigma^2} \right) }
\end{eqnarray}

From a statistical viewpoint, the multiplicative nature of amplitude and intensity data can be explained as follows. 
Let us consider two fixed, independent to $\sigma$ unit distributions given below:

\begin{eqnarray}
pdf(A_1) &=& 2A_1 e^{ \left( A_1^2 \right) }\\
pdf(I_1) &=& e^{ \left( -I_1 \right) }
\end{eqnarray}

Using the variable change theorem,
%  it is then trivial to show that:
    it can be shown that:
    the SAR amplitude and intensity are equal to %are simply
    a scaled version of these unit variables,
    specifically $A= \sigma A_1 $ and $I= \sigma^2 I_1 $. 
These relationships evidently manifest a mutiplicative nature. 
In fact, the condition has long been noted,
  but from different perspectives,
  in various proposed SAR models (e.g. \cite{Jakeman_1980_JPhysAMathGen}).

Let us now consder the population expected mean and variance of the four distributions.
They are given in Table \ref{tbl:orginal_sar_avg_var}. 
If spatial homogeneity is defined as an imaging areas having the same back-scattering coefficient $\sigma$,
  then over a homogeneous scene, the measured values can then be considered as samples coming from a single stochastic process.
However, over heterogeneous areas where $\sigma$ varies significantly,
  it is then evidently clear that: the amplitude ($A$) as well as the intensity ($I$) of SAR data suffer from hetoroskedastic phenomena,
  which is defined as the dependence of conditional expected variance of SAR data on the conditional expectation of mean. 
In the context of speckle filtering which aims to estimate the unknown parameter $\sigma$,
  Table \ref{tbl:orginal_sar_avg_var} also clearly indicates the ambiguous circle.
That is, in the original domain, estimating variance is equal to estimating the mean
  and is as hard as the main problem itself.
  %, i.e. estimating the unknown parameter $\sigma$

@@

The formulas above have long been noted in this research field. 
In fact while pioneering the estimation of the equivalent number of look (ENL) index, 
 it has been noted that the ratio of expected standard deviation to mean is a constant in both cases (i.e. $snr(A)=\sqrt{\frac{4}{\pi}-1}$ and $snr(I)=1$) \cite{Lee_1992_IMA_298}.
Here we just offer a different perspective,
  one that avoids the negative effects of heteroskedasticity in SAR data processing.
%Here, we just offer a different interpretation, which sets the stage for the discussion of logarithmic transformation.

\section{The Effects of Heteroskedasticity in SAR Data Processing}

@@

The nature of SAR speckle is stochastic over homogeneous areas and heteroskedastic heterogeneously. 
Statistical models are used to describe how the underlying back-scattering coefficient ($\sigma$) affects the distribution of the observable SAR data. 
%For about 30 years, speckle filtering has been an active research area, with new methods being introduced steadily \cite{Lee_TGRS_2009}. 
%That is because
And while the statistical model within individual resolution cells is well established,
  the applicability of these models, however, are restricted to homogeneous areas. 
Practical images, nevertheless, are heterogeneous. 
Crucially it is this spatial variation that is of high interest. 
This fact gives rise to the question that seemed obvious:
  %how to call an analysis area heterogeneous and subsequently
  what to do in the case of heterogeneity.

Various statistical models for heterogeneous areas have been proposed (see \cite{Touzi_2002_TGRS} for a detailed review). 
Unfortunately, while most of the models highlight the multiplicative nature of sub-pixel or homogeneous original SAR data, in extending the model to heterogeneous images, virtually none has noted that spatial variation also gives rise to heteroskedasticity phenomena. 
Heteroskedasticity, as explained later, is defined as the dependence of conditional expected variance of original SAR data on the conditional expectation of its mean or equivalently its underlying back-scattering coefficient $(\sigma)$.

Speckle filtering, by and large, are estimators attempting to estimate the unknown coefficient $\sigma$ from the observable SAR signal $(A)$. 
Thus, SAR speckle-filtering can be and has been positioned within the context of estimation theory \cite{Touzi_2002_TGRS}. 
The statistical framework attempts to estimate unknown statistical parameters from the observable data. 
The stages consist of statistical modelling, estimator development and evaluating the performance of the statistical estimators. 
%In this section, the negative effects of heteroskedasticity on speckle filtering is analysed.
It is believed that heteroskedasticity, in its turn, give rise to serious negative impacts on various stages of speckle-filtering.

In modelling, heteroskedasticity has direct consequences in the central question of homogeneity or heterogeneity. 
In optical images, the contrast or variance among neighboring pixels have been used to measure homogeneity. 
Unfortunately such techniques do not appear to be effective under the heteroskedastic condition of original SAR data. 
In SAR images, both of these measures are dependent on the underlying signal $(\sigma)$. 
This make the problem of estimating variance equal to the problem of estimating the mean and is the same as the main problem: estimating $\sigma$. 
The fact give rise to the ambiguity circle in SAR speckle filtering.

A case in point can be found in the recently published improved sigma filter\cite{Lee_TGRS_2009}. 
The technique determines outlying points as being too far away from the standard deviation. 
However, as is done in \cite{Lee_TGRS_2009}, to estimate standard deviation, an estimator of the mean is required and used. 
It is interesting to note that the MMSE estimator used to estimate the mean in \cite{Lee_TGRS_2009}, itself alone, is a rather successful speckle-filter \cite{Lee_PAMI_1980}.

Heteroskedasticity also poses numerous challenges in designing an efficient estimator. Heteroskedasticity directly violates Gauss-Markov theorem's homoskedastic assumption. 
Thus it renders the efficiency of any naive Ordinary Least Square estimator in serious doubt. 
If the variance is known \textit{a priori}, it has been proven \cite{Aitken_1934_ProcsRoyalSocEdinburg}, that a weighted mean estimator is the best linear unbiased estimator. 
Interestingly, as noted by Lopes \cite{Lopes_TGRS_1990}, most known common successful adaptive filters \cite{Lee_PAMI_1980, Kuan_1985_PAMI, Frost_PAMI_1982} do make use of weighted mean estimations. 
The caveat however is that in SAR speckle filtering, variance is not known \textit{a priori}. 
And even though variance can be estimated from observable values, as the ambiguity circle goes, estimating the variance is as hard as estimating the underlying coefficient $\sigma$ itself.

Heteroskedasticity also affects the performance evaluation stage of such a framework.
Statistical estimators are routinely evaluated using their bias and variance performance,
  which when combined normally results in MSE evaluation.
Unfortunately, under heteroskedastic conditions, Least Squared Error estimation may not be optimal.   
In other words, Minimal Squared Error can no longer serves as a reliable evaluation metric.
A number of different evaluation metrics have been proposed for SAR speckle filters,
  however, none have appeared capable of taking on the MSE's previous role.

%Heteroskedasticity also decrease the effectiveness of estimators by increasing the variance of even the best possible estimations. 
%This fact makes it tricky to evaluate the performance of different estimators, even if quantitative experimental results are reported.
%In fact several quantitative simulated experiments has been carried out to compare the performance of different estimators\cite{Lee_TGRS_2009}. 
%However only results of single experiments were reported. 
%Due to the stochastic nature of the simulated experiment process, more experiments should be done to report the effectiveness, i.e. variance, of each estimators' results.

Last but certainly not least, is the bad impact of heteroskedasticity on SAR image interpretation. 
Most of the task to be carried out in interpreting SAR images almost invariably involve target detection, target segmentation and/or target classification. 
All of these tasks require good discriminant functions. 
Foundational to all of these is the need for a consistence sense of distance. 
Unfortunately, by definition, heteroskedasticity provides inconsistent measures of distance. 
This inconsistent sense of distance coupled with the failure of the ordinary least square regression methods are believed to have caused a large class of artificial neural networks as well as a number of other computational intelligence methods to under-perform on SAR data.

%In our work, we aim to neutralize these various negative impacts. 
%The paper is organized as follows. 
%Section \ref{sec:heteroskdastic_model} will be devoted to discussion of the heteroskedastic phenomena in the original per-pixel SAR amplitude and intensity model. 
%Logarithmic transformation is extended to the model and the homoskedastic property of it is asserted. 
%Section \ref{sec:variance_in_log_domain} will discuss how sampling distribution of variance in the log-transformed domain can be used to model various degrees of homogeneity. 
%Section \ref{sec:k_mle} will presents a novel speckle filtering algorithm developed from heterogenous inference made possible by statistical testing of sample variances.
%Section \ref{sec:k_mle_result} will present evaluation of our estimator both quantitatively through simulated experiments and qualitatively on real-life images.

@@

%\subsection{Homoskedastic effect of Logarithmic Transformation }

In this chapter, a base-2 logarithmic transformtion is proposed for SAR data. 
The choice of base-2 is due to implementation reasons
  since base-2 logarithm can be computed faster on digital computer than either natural or decimal logarithms,
  and yet the ability to transform heterskedactic speckle into a homoskedastic relationship is still maintained.
Under base-2 log-transformation, the original variables become:

\begin{eqnarray}
L_{1}^{A} &=& \log_2(A_1) = L_{1}^{I} / 2 \\
L_A &=& \log_2(A) 	= L_{1}^{A} + \log_2\sigma \\
L_{1}^{I} &=& \log_2(I_1) = 2 L_{1}^{A} \\
L_I &=& \log_2(I) 	= L_{1}^{I} + 2 \log_2\sigma
\end{eqnarray}

The equations above highlight the relationships among random variables in the log-transformed domain. 
Two conclusions become evident from these formulae. 
Firstly, in the log-transformed domain, working on either amplitude or intensity will tend to give identical results. 
Secondly, the effects of converting the multiplicative nature to an additive nature through logarithmic transformation
is clearly manifested. 
Now, bearing the relationship among the random variables in mind, it is then trivial to 
show that the probability distribution of these log-transformed variables are related as follows:

\begin{eqnarray}
pdf(L_{1}^{A}) &=& 2 \cdot 2^{\left( 2 L_{1}^{A} \right)}e^{-2^{\left( 2 L_{1}^{A} \right) }} \\
pdf(L_A) &=& 2 \cdot 2^{\left( 2 L_A - 2 \log_2 \sigma \right)}e^{-2^{\left( 2 L_A - 2 \log_2 \sigma \right)}} \\ 
pdf(L_{1}^{I}) &=& 2^{L_1^I} e^{-2^{L_1^I}}  \\
pdf(L_I) &=& 2^{\left( L_I - 2 \log_2 \sigma \right)} e^{-2^{\left( L_I - 2 \log_2 \sigma \right)} }  
%pdf(L_{1}^{A}) &=& 2 \cdot 2^{\left[ 2 L_{1}^{A} - 2^{2 L_{1}^{A}} \right]} \\
%pdf(L_A) &=& 2 \cdot 2^{\left[ \left( 2 L_A - 2 \log_2 \sigma \right) - 2^{\left( 2 L_A - 2 \log_2 \sigma \right)} \right]} \\ 
%pdf(L_{1}^{I}) &=& 2^{\left[ L_{1}^{I} - 2^{L_{1}^{I}} \right]} \\
%pdf(L_I) &=& 2^{\left[ \left( L_I - 2 \log_2 \sigma \right) - 2^{\left( L_I - 2 \log_2 \sigma \right)} \right]} \label{eqn:log_intensity_base_2_dist}
\end{eqnarray}

Noting that these distributions belong to the Fisher-Tippet family, the population expected mean and variances 
are obtained and presented in Table \ref{tbl:sar_log_domain_avg_var}, with $\gamma \approx 0.577$ being the Euler-Mascheroni constant. 
Most importantly, it can be seen that the means are biased and the variances are no longer related to $\sigma$

@@

Table \ref{tbl:sar_variables_properties} summarizes the discussion so far. 
It can be seen that while the original data, especially intensity values, should be preferred for 
multi-look processing, the log-transformed domain with its homoskedastic distribution offers consistent sense of variance.

@@

To experimentally validate this theoretical discussion, Fig. \ref{fig:modelled_response} plots the histogram of observable data within a 
known homogenous area (from a RADARSAT2 image) against modelled PDF response. 
An excellent agreement is observable in the graph which validate the log-transformed model. 
%In fact, this modelling has been used successful in explaining speckle phenomena, also verified 
%through scientific experiments \cite{Ulaby_TGRS_1988}.

\begin{figure}[!h]
\centering
%\centerline{
\begin{tabular}{c}
	\subfloat[amplitude]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/amplitude_histogram.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[intensity]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/intensity_histogram.eps} 	
		 \label{intensity}
	} \\
	\subfloat[log amplitude]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/log_amplitude_histogram.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[log intensity]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/log_intensity_histogram.eps} 	
		 \label{intensity}
	} 
\end{tabular}
%}
\caption{Observed histogram in homogenous area and modelled pdf response}
\label{fig:modelled_response}
\end{figure}

@@

%\subsection{Consistent measures of distance in Log-Transformed domain}

The consistent feature is explored and illustrated in this section from two different perspectives. 
First, the back-scattering coefficient $\sigma$ is assumed to be \textit{known a priori}.
With that, let us consider the dispersion random 
variable defined as the distance between an observable sample and its expected value:

\begin{eqnarray}
D_A &=& A - avg(A) \\
D_I &=& I - avg(I) \\
D_{L^A} &=& L_A - avg(L_A) \\
D_{L^I} &=& L_I - avg(L_I) 
\end{eqnarray}

Noting the results from the previous section, the PDF for these variable can be derived as follows:

\begin{eqnarray}
pdf(D_A) &=& 2 \cdot \frac{\left( D_A + \sigma \sqrt{\pi}/2 \right)}{\sigma^2}e^{ \left[ - \frac{\left( D_A + \sigma \sqrt{\pi}/2 \right)^2}{\sigma^2}   \right] } \\
pdf(D_I) &=& \frac{1}{\sigma^2}e^{\left[ -\left( D_I + \sigma^2 \right) / \sigma^2 \right] } \\
pdf(D_{L^A}) &=& 2 \cdot 2^{\left( 2 D_{L^A} + 2 \frac{\gamma}{2 \ln2} \right)} e^{-2^{\left( 2 D_{L^A} + 2 \frac{\gamma}{2 \ln2} \right)}}\\
pdf(D_{L^I}) &=& 2^{\left( D_{L^I} + \frac{\gamma}{\ln2} \right)} e^{-2^{\left( D_{L^I} + \frac{\gamma}{\ln2} \right)}}
%pdf(D_{L^A}) &=& 2 \cdot 2^{\left[ \left( 2 D_{L^A} + 2 \frac{\gamma}{2 \ln2} \right) - 2^{\left( 2 D_{L^A} + 2 \frac{\gamma}{2 \ln2} \right)} \right]} \\
%pdf(D_{L^I}) &=& 2^{\left[ \left( D_{L^I} + \frac{\gamma}{\ln2} \right) - 2^{\left( D_{L^I} + \frac{\gamma}{\ln2} \right)} \right]}
\end{eqnarray}

From these equations, it is clear that these distributions, which are dependent on $\sigma$ in the original domain, 
are now independent of it when processed in the log-domain.

Let us investigate SAR data from the second perspective,
  where two adjacent resolution cells are known to have \textit{an identical but unknown} back-scattering coefficient $\sigma$.
Now, the contrast random variable is defined, also in the log-transformed domain, as the distance between two measured samples.

%From a second perspective, given two adjacent resolution cells that are known to have \textit{identical but unknown} back-scattering coefficient $\sigma$,
%  consider the contrast random variable defined, also in the log-transformed domain, as the distance between two 
%measured samples:

\begin{eqnarray}
C_A &=& A_1^\sigma - A_2^\sigma \\
C_I &=& I_1^\sigma - I_2^\sigma \\
C_{L^A} &=& L_1^{A_\sigma} - L_2^{A_\sigma} = log_2 { \left( {A_1}/{A_2} \right) }\\
C_{L^I} &=& L_1^{I_\sigma} - L_2^{I_\sigma} = log_2 { \left( {I_1}/{I_2} \right) }
\end{eqnarray}

Noting that $C_x = D_1^x - D_2^x$, it should come as no suprise that the measure of contrast is also consistent in the 
log-domain but inconsistent in the original domain. The PDF of the log-transformed variables can be expressed as:

>>>>

\chapter{Homoskedastic Models for Heteroskedastic SAR Data} %chapter 3
\label{chap:sar}

This chapter  highlights that SAR
                data are heteroskedastic in the original domain.
Relying solely on statistical and mathematical principles,             
the log-transformed SAR data is shown to be homoskedastic.
From there, the consistent sense of variance and a few consistent measures of distance are proposed and validated against real-life
                practical SAR data.

@@

The stochastic nature of SAR data has been illustrated earlier (see Eqns. \ref{eqn:SAR Amplitude PDF} and \ref{eqn:SAR Intensity PDF}).
From a statistical viewpoint, the multiplicative nature of amplitude ($A$) and intensity ($I$)  data can be explained as follows. 
Let us consider two fixed unit distributions which are independent to $\sigma$ given below:

\begin{equation}
\label{eqn:SAR Unit Amplitude PDF}
\caption{eqn:SAR Unit Amplitude PDF}
pdf(A_1) = 2A_1 e^{ \left( A_1^2 \right) }
\end{equation}
\begin{equation}
  \label{eqn:SAR Unit Intensity PDF}
  \caption{eqn:SAR Unit Intensity PDF}
pdf(I_1) = e^{ \left( -I_1 \right) }
\end{equation}

Using the variable change theorem,
    it can be shown that
    the SAR amplitude and intensity are equal to %are simply
    a scaled version of these unit variables,
    specifically $A= \sigma A_1 $ and $I= \sigma^2 I_1 $. 
These relationships evidently display a mutiplicative nature. 
In fact, the condition has long been noted,
  but from different perspectives,
  in various proposed SAR models (e.g. \cite{Jakeman_1980_JPhysAMathGen}).

Let us now consider the population expected mean and variance of the four statistical distributions for these observables, i.e. $A, A_1, I, I_1$.
They are given in Table \ref{tbl:orginal_sar_avg_var}. 
If spatial homogeneity is defined as an imaging area having the same back-scattering coefficient $\sigma$,
  then over a homogeneous scene, the measured values can then be considered as samples coming from a single stochastic process.
However, over heterogeneous areas where $\sigma$ varies significantly,
  it is then evidently clear that: the amplitude  as well as the intensity of SAR data suffer from hetoroskedastic phenomena. 
%  which is defined as the dependence of conditional expected variance of SAR data on the conditional expectation of mean.
This means that  the conditional expected variance of SAR data is dependent on the conditional expectation of the mean.   
In the context of speckle filtering which aims to estimate the unknown parameter $\sigma$,
  Table \ref{tbl:orginal_sar_avg_var} also clearly indicates an ambiguous circle:
in the original domain, estimating variance is equal to estimating the mean, but this is as hard to estimation as the variance.

@@

The formulas above are extremely well known in this research field. 
In fact while pioneering the estimation of the equivalent number of look (ENL) index, 
 it was noted by Lee et.al. that the ratio of expected standard deviation to mean is a constant in both cases (i.e. $snr(A)=\sqrt{\frac{4}{\pi}-1}$ and $snr(I)=1$) \cite{Lee_1992_IMA_298}.
%Here we just offer a different perspective,
%  one that avoids the negative effects of heteroskedasticity in SAR data processing.
%Here, we just offer a different interpretation, which sets the stage for the discussion of logarithmic transformation.

In this thesis, a different perspective will be proposed which avoids the negative effects of heteroskedasticity in SAR data processing.

\section{The Effects of Heteroskedasticity in SAR Processing}%IVM: shortened to make it look nicer

@@

The nature of SAR speckle is stochastic over homogeneous areas and heteroskedastic heterogeneously. 
Statistical models are used to describe how the underlying back-scattering coefficient ($\sigma$) affects the distribution of the observable SAR data. 
While the statistical model within individual resolution cells is well established,
  it is only applicable to homogeneous areas.
Practical images, however, are heterogeneous.
Furthermore, it is precisely this spatial variation that is often of high interest.
%And while the statistical model within individual resolution cells is well established,
%  the applicability of these models, however, are restricted to homogeneous areas. 
%Practical images, nevertheless, are heterogeneous. 
%Crucially it is this spatial variation that is of high interest. 
%This fact gives rise to the question that seemed obvious:
  %how to call an analysis area heterogeneous and subsequently
%  what to do in the case of heterogeneity.

As such, various statistical models for heterogeneous areas have been proposed (see \cite{Touzi_2002_TGRS} for a detailed review). 
However, while most of the models highlight the multiplicative nature of sub-pixel or homogeneous original SAR data, in extending the model to heterogeneous images, virtually none have noted that spatial variation also gives rise to heteroskedasticity phenomena. 
Heteroskedasticity, as explained later, is defined as the dependence of conditional expected variance of original SAR data on the conditional expectation of its mean or equivalently its underlying back-scattering coefficient $(\sigma)$.

Speckle filters are effectively estimators attempting to estimate the unknown coefficient $\sigma$ from the observable SAR signal $(A)$. 
Thus, SAR speckle-filtering can be and has been positioned within the context of estimation theory \cite{Touzi_2002_TGRS}, where
the statistical framework is used to estimate unknown statistical parameters from the observable data. 
The stages of this framework consist of statistical modelling, estimator development and evaluating the performance of the statistical estimators. 
%In this section, the negative effects of heteroskedasticity on speckle filtering is analysed.
It is also known that heteroskedasticity gives rise to serious negative impacts on various stages of speckle-filtering.

In modelling, heteroskedasticity has direct consequences in the central question of homogeneity or heterogeneity. 
In optical images, the contrast or variance among neighbouring pixels have been used to measure homogeneity. 
Unfortunately such techniques do not appear to be effective under the heteroskedastic condition of original SAR data. 
In SAR images, both of these measures are dependent on the underlying signal $(\sigma)$. 
This make the problem of estimating variance equal to the problem of estimating the mean and is the same as the main problem: estimating $\sigma$. 
The fact give rise to the ambiguity circle in SAR speckle filtering.

A case in point can be found in a recently published improved sigma filter \cite{Lee_TGRS_2009}. 
The technique determines outlying points as being too far away from the standard deviation. 
However, as is done in \cite{Lee_TGRS_2009}, to estimate standard deviation, an estimator of the mean is required and used. 
It is interesting to note that the MMSE estimator used to estimate the mean in \cite{Lee_TGRS_2009}, itself alone, is a rather successful speckle-filter \cite{Lee_PAMI_1980}.

Heteroskedasticity also poses numerous challenges in designing an efficient estimator. Heteroskedasticity directly violates Gauss-Markov theorem's homoskedastic assumption. 
Thus it renders the efficiency of any naive Ordinary Least Square estimator in serious doubt. 
If the variance is known \textit{a priori}, it has been proven \cite{Aitken_1934_ProcsRoyalSocEdinburg}, that a weighted mean estimator is the best linear unbiased estimator. 
Interestingly, as noted by Lopes \cite{Lopes_TGRS_1990}, most known common successful adaptive filters \cite{Lee_PAMI_1980, Kuan_1985_PAMI, Frost_PAMI_1982} do make use of weighted mean estimations. 
The problem, however, is that in SAR speckle filtering, variance is not known \textit{a priori}. 
And even though variance can be estimated from observable values, as the ambiguity circle goes, estimating the variance is as hard as estimating the underlying coefficient $\sigma$ itself.

Heteroskedasticity also affects the performance evaluation stage of the estimation framework.
Statistical estimators are routinely evaluated using their bias and variance performance
  which, when combined, normally result in MSE evaluation.
Unfortunately, under heteroskedastic conditions, Least Squared Error estimation may not be optimal.   
In other words, Minimum Squared Error can no longer serves as a reliable evaluation metric.
A number of different evaluation metrics have been proposed for SAR speckle filters,
  however, none have appeared capable of taking on the MSE's previous role.
%IVM: Hai, the above "Minimum Squared Error" should be Mean Squared Error???

Last but certainly not least, is the bad impact of heteroskedasticity on SAR image interpretation. 
Most of the task to be carried out in interpreting SAR images almost invariably involve target detection, target segmentation and/or target classification. 
All of these tasks require good discriminant functions. 
Fundamental to all of those is the need for a consistent measure of distance. 
Unfortunately, by definition, heteroskedasticity provides inconsistent measures of distance. 
This inconsistent measure of distance coupled with the failure of the ordinary least square regression methods have caused a large class of artificial neural networks as well as a number of other computational intelligence methods to under-perform for SAR data.

@@

In this section, a base-2 logarithmic transformation is proposed for SAR data. 
The choice of base-2 is due to implementation reasons,
  since base-2 logarithms can be computed faster by computer than either natural or decimal logarithms,
  and yet the ability to transform heterskedactic speckle into a homoskedastic relationship is still maintained.
Under base-2 log-transformation, the original variables become:

\begin{equation}
  \label{eqn:SAR Unit Log Amplitude Observable}
  \caption{eqn:SAR Unit Log Amplitude Observable}
L_{1}^{A} = \log_2(A_1) = L_{1}^{I} / 2 
\end{equation}
\begin{equation}
  \label{eqn:SAR Log Amplitude Observable}
  \caption{eqn:SAR Log Amplitude Observable}
L_A = \log_2(A) 	= L_{1}^{A} + \log_2\sigma 
\end{equation}
\begin{equation}
  \label{eqn:SAR Unit Log Intensity Observable}
  \caption{eqn:SAR Unit Log Intensity Observable}
L_{1}^{I} = \log_2(I_1) = 2 L_{1}^{A} 
\end{equation}
\begin{equation}
  \label{eqn:SAR Log Intensity Observable}
  \caption{eqn:SAR Log Intensity Observable}
L_I = \log_2(I) 	= L_{1}^{I} + 2 \log_2\sigma
\end{equation}

The equations above highlight the relationships among random variables in the log-transformed domain. 
Two interesting points are evident from these formulae. 
Firstly, in the log-transformed domain, working on either amplitude or intensity will tend to give identical results. 
Secondly, the effects of converting the multiplicative nature to an additive nature through logarithmic transformation
is clearly manifested.
Now, bearing in mind the relationship among the random variables, %it is then trivial to 
%show that the probability distribution of these log-transformed variables are related as follows:
application of variable change theorem leads to:

\begin{equation}
  \label{eqn:SAR Unit Log Amplitude PDF}
  \caption{eqn:SAR Unit Log Amplitude PDF}
pdf(L_{1}^{A}) = 2 \cdot 2^{\left( 2 L_{1}^{A} \right)}e^{-2^{\left( 2 L_{1}^{A} \right) }} 
\end{equation}
\begin{equation}
  \label{eqn:SAR Log Amplitude PDF}
  \caption{eqn:SAR Log Amplitude PDF}
pdf(L_A) = 2 \cdot 2^{\left( 2 L_A - 2 \log_2 \sigma \right)}e^{-2^{\left( 2 L_A - 2 \log_2 \sigma \right)}} 
\end{equation}
\begin{equation}
  \label{eqn:SAR Unit Log Intensity PDF}
  \caption{eqn:SAR Unit Log Intensity PDF}
pdf(L_{1}^{I}) = 2^{L_1^I} e^{-2^{L_1^I}}  
\end{equation}
\begin{equation}
  \label{eqn:SAR Log Intensity PDF}
  \caption{eqn:SAR Log Intensity PDF}
pdf(L_I) = 2^{\left( L_I - 2 \log_2 \sigma \right)} e^{-2^{\left( L_I - 2 \log_2 \sigma \right)} }  
\end{equation}

Noting that these distributions belong to the Fisher-Tippet family, the population expected mean and variances 
are obtained and presented in Table \ref{tbl:sar_log_domain_avg_var}, with $\gamma \approx 0.577$ being the Euler-Mascheroni constant. 
Most importantly, it can be seen that the means are biased and the variances are no longer related to $\sigma$.

@@

The discussion so far is further summarized in Table \ref{tbl:sar_variables_properties}.
It shows that while the original data, especially intensity values, should be preferred for 
multi-look processing, the log-transformed domain with its homoskedastic distribution offers a consistent sense of variance.

@@

To experimentally validate this theoretical discussion, real world data is extracted from a known  homogenous area (from within a RADARSAT2 image). The PDF histogram from this data is then plotted in 
Fig. \ref{fig:modelled_response} along with the modelled PDF response. 
The excellent agreement observable between the real world and modelled data tends to validate the log-transformed model. 
%In fact, this modelling has been used successful in explaining speckle phenomena, also verified 
%through scientific experiments \cite{Ulaby_TGRS_1988}.

\begin{figure}[!h]
\centering
%\centerline{
\begin{tabular}{c}
	\subfloat[amplitude]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/amplitude_histogram.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[intensity]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/intensity_histogram.eps} 	
		 \label{intensity}
	} \\
	\subfloat[log amplitude]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/log_amplitude_histogram.eps} 	
		 \label{amplitude}
	} 
	\hfill	
	\subfloat[log intensity]{
		 \epsfxsize=2.5in
		 \epsfysize=2.5in
		 \epsffile{images/log_intensity_histogram.eps} 	
		 \label{intensity}
	} 
\end{tabular}
%}
\caption{Observed histogram of a homogenous area and modelled PDF response}
\label{fig:modelled_response}
\end{figure}

@@

The consistent feature is explored and illustrated in this section from two different perspectives. 
First, the back-scattering coefficient $\sigma$ is assumed to be \textit{known a priori}.
With that, let us consider the dispersion random 
variable which is defined as the distance between an observable sample and its expected value:

\begin{equation}
  \label{eqn:SAR Amplitude Distance Observable}
  \caption{eqn:SAR Amplitude Distance Observable}
D_A = A - avg(A)   
\end{equation}
\begin{equation}
  \label{eqn:SAR Intensity Distance Observable}
  \caption{eqn:SAR Intensity Distance Observable}
D_I = I - avg(I)  
\end{equation}
\begin{equation}
  \label{eqn:SAR Log Amplitude Distance Observable}
  \caption{eqn:SAR Log Amplitude Distance Observable}
D_{L^A} = L_A - avg(L_A)
\end{equation}
\begin{equation}
  \label{eqn:SAR Log Intensity Distance Observable}
  \caption{eqn:SAR Log Intensity Distance Observable}
D_{L^I} = L_I - avg(L_I) 
\end{equation}

Noting the results from the previous section, the PDF for these variables can be derived as follows:

\begin{equation}
  \label{eqn:SAR Amplitude Distance PDF}
  \caption{eqn:SAR Amplitude Distance PDF}
pdf(D_A) = 2 \cdot \frac{\left( D_A + \sigma \sqrt{\pi}/2 \right)}{\sigma^2}e^{ \left[ - \frac{\left( D_A + \sigma \sqrt{\pi}/2 \right)^2}{\sigma^2}   \right] }
\end{equation}
\begin{equation}
  \label{eqn:SAR Intensity Distance PDF}
  \caption{eqn:SAR Intensity Distance PDF}
pdf(D_I) = \frac{1}{\sigma^2}e^{\left[ -\left( D_I + \sigma^2 \right) / \sigma^2 \right] } \\
\end{equation}
\begin{equation}
  \label{eqn:SAR Log Amplitude Distance PDF}
  \caption{eqn:SAR Log Amplitude Distance PDF}
pdf(D_{L^A}) = 2 \cdot 2^{\left( 2 D_{L^A} + 2 \frac{\gamma}{2 \ln2} \right)} e^{-2^{\left( 2 D_{L^A} + 2 \frac{\gamma}{2 \ln2} \right)}}
\end{equation}
\begin{equation}
  \label{eqn:SAR Log Intensity Distance PDF}
  \caption{eqn:SAR Log Intensity Distance PDF}
pdf(D_{L^I}) = 2^{\left( D_{L^I} + \frac{\gamma}{\ln2} \right)} e^{-2^{\left( D_{L^I} + \frac{\gamma}{\ln2} \right)}}
\end{equation}

It is clear that these distributions are dependent on $\sigma$ in the original domain (the upper two equations), are now independent of it when processed in the log-domain (the lower two equations).

Let us now investigate SAR data from the second perspective,
  where two adjacent resolution cells are known to have \textit{an identical but unknown} back-scattering coefficient $\sigma$.
Now, the contrast random variable is defined, also in the log-transformed domain, as the distance between two measured samples.

\begin{equation}
  \label{eqn:SAR Amplitude Contrast Observable}
  \caption{eqn:SAR Amplitude Contrast Observable}
C_A = A_1^\sigma - A_2^\sigma 
\end{equation}
\begin{equation}
  \label{eqn:SAR Intensity Contrast Observable}
  \caption{eqn:SAR Intensity Contrast Observable}
C_I = I_1^\sigma - I_2^\sigma 
\end{equation}
\begin{equation}
  \label{eqn:SAR Log Amplitude Contrast Observable}
  \caption{eqn:SAR Log Amplitude Contrast Observable}
C_{L^A} = L_1^{A_\sigma} - L_2^{A_\sigma} = log_2 { \left( {A_1}/{A_2} \right) }
\end{equation}
\begin{equation}
  \label{eqn:SAR Log Intensity Contrast Observable}
  \caption{eqn:SAR Log Intensity Contrast Observable}
C_{L^I} = L_1^{I_\sigma} - L_2^{I_\sigma} = log_2 { \left( {I_1}/{I_2} \right) }
\end{equation}
